---
title: "coda4microbiome-oc_pub"
author: "Fernanda Terrazas"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
Set working directory
```{r}
# setwd("~/Documents/Programming/Microbiome/microbiome-ocp/data-analyses")

```
Load coda4microbiome package
```{r Load Packages}
library(coda4microbiome)

```

```{r}
# to reproduce the results
set.seed(123)

```
# 5. coda4microbiome - CH4 & H2 levels signatures in Control/Users
Assign "group" with metadata subset
```{r}
# group <- map[!(map$SampleID %in% remove_outliers),]
# group <- map[(map$SampleID %in% merged_hydro_meth_auc_map$SampleID),]

# add H2 and CH4 AUC values to map data (use hydro_meth_df)
group <- merge(map, hydro_meth_df[c("SampleID", "H2", "CH4")], by = "SampleID")
head(group)

```
### Use group and complete.cases() to save CH4 and H2 columns that have data for all entries, then convert responses to binary
```{r}
### Convert breath AUC data to binary (high or low)
### Calculate mean and categorize responses as high or low (above or below the mean)
y_hydrogen <- subset(group, complete.cases(H2))
# y_hydrogen$H2 <- sqrt(y_hydrogen$H2)
mean_H2 <- mean(y_hydrogen$H2) # calculate the mean of H2
y_hydrogen$b_H2 <- ifelse(y_hydrogen$H2 > mean_H2, "high", "low") # create a b_H2 column with high/low values for H2
y_hydrogen_b <- y_hydrogen$b_H2

y_methane <- subset(group, complete.cases(CH4))
# y_methane$CH4 <- sqrt(y_methane$CH4)
mean_CH4 <- mean(y_methane$CH4) # calculate the mean of CH4
y_methane$b_CH4 <- ifelse(y_methane$CH4 > mean_CH4, "high", "low") # create a b_CH4 column with high/low values for CH4
y_methane_b <- y_methane$b_CH4

```
### Run coda4microbiome algorithm
##### **Hydrogen** AUC values between Control and Users
```{r}
### Balances for Hydrogen Level Differences

### 1. Create a Transposed Data Frame. A transposed data frame x_species is created from the original data frame species.
x_species <- data.frame(t(species), check.names = T)
# subset asv/species/genus data to keep subgroup data
# x_species <- x_species[row.names(x_species) %in% group$SampleID, ]
x_species_hydro <- x_species[row.names(x_species) %in% y_hydrogen$SampleID, ]

### 2. Factorize the Low/High levels of Hydrogen. Create a factor variable cu_vector_hydro by extracting the hydrogen information for 'Control' & 'User' subjects from the y_hydrogen(from group, from map) data.
# 'Control' samples only
# y_control_vector_hydro <- factor(y_hydrogen[(y_hydrogen$Treatment=="Control" ),]$b_H2)
# 'User' samples only
# y_user_vector_hydro <- factor(y_hydrogen[(y_hydrogen$Treatment=="User" ),]$b_H2)
# All samples
y_cu_vector_hydro <- factor(y_hydrogen$b_H2)

### 3. Apply regression. Regression is applied using the coda_glmnet function to identify important taxa in explaining H2 level differences in 'Control' & 'User' subjects.
# only take taxa over 10% contribution to variance
coda_glmnet_hydro <- coda_glmnet(x = x_species_hydro, y = y_cu_vector_hydro, lambda = "lambda.min", nvar = NULL, alpha = 0.9, nfolds = 7, showPlots = TRUE, coef_threshold = 0)

### 4. Extract Accuracy Measure. The apparent Area Under the Curve (AUC) is extracted as an accuracy measure from the regression results.
coda_glmnet_hydro$`apparent AUC`
coda_glmnet_hydro$`predictions plot`
coda_glmnet_hydro$`signature plot`

### 5. Create a Data Frame with Genera and Coefficients. A data frame coda_balance_coeff_hydro is created to store genera names and their corresponding coefficients from the regression
coda_balance_coeff_hydro <- data.frame(cbind(coda_glmnet_hydro$`log-contrast coefficients`, coda_glmnet_hydro$taxa.name))

### 6. Data Frame Manipulations
# Set row names of the data frame coda_balance_coeff_hydroto be the taxa names obtained from the regression results
rownames(coda_balance_coeff_hydro) <- coda_balance_coeff_hydro$taxa.name
# Set column names of the data frame to "Coefficient_full" and "Species" for clarity.
colnames(coda_balance_coeff_hydro) <- c("Coefficient_full", "Species")
# Convert "Coefficient_full" column to numeric type, and create a new column "Coefficient" by rounding "Coefficient_full" to two decimal places
coda_balance_coeff_hydro$Coefficient_full <- as.numeric(coda_balance_coeff_hydro$Coefficient_full)
coda_balance_coeff_hydro$Coefficient <-as.numeric(format(round(coda_balance_coeff_hydro$Coefficient_full, 2), nsmall = 2))
# Determine Positive or Negative Coefficients. Create "PosNeg" column based on whether the coefficient is positive or negative; contains the labels "Positive" or "Negative" accordingly.
coda_balance_coeff_hydro$PosNeg <- with(coda_balance_coeff_hydro, ifelse(Coefficient > 0, "Positive", "Negative"))
# Assign Group Labels. The "Group" column is assigned labels "High" or "Low" based on the "PosNeg" column. If the coefficient is positive, it is labeled as "Low"; otherwise, it is labeled as "High".
coda_balance_coeff_hydro$Group <- with(coda_balance_coeff_hydro, ifelse(PosNeg == "Positive", "Low", "High"))

### 7. Create Data Frame with Samples, Predictions, and Variables. A data frame coda_balances_hydro_predictions is created to store balance predictions along with sex, group, and sample type information
coda_balances_hydro_predictions <- data.frame(coda_glmnet_hydro$predictions)
colnames(coda_balances_hydro_predictions) <- c("Balance")
coda_balances_hydro_predictions$H2_level <- y_cu_vector_hydro
# coda_balances_hydro_predictions$Treatment <- factor(group[(group$Treatment=="Control"),]$Treatment)
coda_balances_hydro_predictions$Treatment <- factor(group$Treatment)
# coda_balances_hydro_predictions$Day <- factor(group[(group$Day=="Day 1" ),]$Day)
coda_balances_hydro_predictions$Day <- factor(group$Day)

### 8. Create Plots. The first one is a boxplot and scatter plot of microbial balances based on sex and group. The second one is a bar plot showing genera in the selected balance, filtered based on a coefficient threshold.
# Graph Sample Balances and Species Coefficients (coeff > 0.099)

# Species coefficient plot
#plot only the genera that are at least 10% of the contribution to the balance
coda_balance_coeff_hydro_10p <- coda_balance_coeff_hydro[(coda_balance_coeff_hydro$Coefficient >= 0.099 | coda_balance_coeff_hydro$Coefficient<= -0.099),]
coda_balance_coeff_hydro_10p

####
#### simplify taxa names using simplify_taxa-names_count-tables.R
####

p_hydro_sample_genera_hydro <- ggplot(
  data = coda_balance_coeff_hydro_10p, 
  aes(x = reorder(Species, Coefficient), y = Coefficient)) +
  geom_bar(stat="identity", width = 0.7, aes(color = Group, fill = Group), show.legend = FALSE) +
  scale_color_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  scale_fill_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  theme_classic() +
  labs(title = "Hydrogen") +
  xlab(" ") + ylab("Log-contrast Coefficient") +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16, hjust = 0.5)) +
  coord_flip() 
p_hydro_sample_genera_hydro

# Balance prediction plot
p_hydro_coda_sample_balances_hydro <- ggplot(data = coda_balances_hydro_predictions, aes(x = H2_level, y = Balance, color = H2_level)) +
    geom_line() +
#    geom_point() +
    geom_boxplot() +
    scale_color_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  scale_fill_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  theme_classic() +
  labs(title = "",
       color = "H2 Level") +
  xlab(" ") + ylab("Balance Value") +
  #facet_grid(~Day) +
    theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size = 16)) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  #theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) +
  coord_flip() +
  theme(legend.position = "none")  # This also removes the legend
p_hydro_coda_sample_balances_hydro


p_coda_glmnet_plots_hydro <- egg::ggarrange(p_hydro_sample_genera_hydro, p_hydro_coda_sample_balances_hydro, labels = c("", ""), 
                         ncol = 1, nrow = 2, heights = c(1,0.25))
as_ggplot(p_coda_glmnet_plots_hydro)

```
Save plot
```{r}
ggsave(filename = "~/Documents/Programming/Microbiome/microbiome-ocp/data-analyses/results/coda4microbiome-pub/p_coda_glmnet_plots_hydro.png", plot = p_coda_glmnet_plots_hydro, width = 5, height = 4, dpi = 300)
```
### Run coda4microbiome algorithm
##### **Methane** AUC values between Control and Users
```{r}
### Balances for methane Level Differences

### 1. Create a Transposed Data Frame. A transposed data frame x_species is created from the original data frame species.
x_species <- data.frame(t(species), check.names = T)
# subset asv/species/genus data to keep subgroup data
# x_species <- x_species[row.names(x_species) %in% group$SampleID, ]
x_species_meth <- x_species[row.names(x_species) %in% y_methane$SampleID, ]

### 2. Factorize the Low/High levels of methane. Create a factor variable cu_vector_methby extracting the methane information for 'Control' & 'User' subjects from the y_methane(from group, from map) data.
# 'Control' samples only
# y_control_vector_meth<- factor(y_methane[(y_methane$Treatment=="Control" ),]$b_CH4)
# 'User' samples only
# y_user_vector_meth<- factor(y_methane[(y_methane$Treatment=="User" ),]$b_CH4)
# All samples
y_cu_vector_meth <- factor(y_methane$b_CH4)

### 3. Apply regression. Regression is applied using the coda_glmnet function to identify important taxa in explaining CH4 level differences in 'Control' & 'User' subjects.
# only take taxa over 10% contribution to variance
coda_glmnet_meth <- coda_glmnet(x = x_species_meth, y = y_cu_vector_meth, lambda = "lambda.min", nvar = NULL, alpha = 0.9, nfolds = 7, showPlots = TRUE, coef_threshold = 0)

### 4. Extract Accuracy Measure. The apparent Area Under the Curve (AUC) is extracted as an accuracy measure from the regression results.
coda_glmnet_meth$`apparent AUC`
coda_glmnet_meth$`predictions plot`
coda_glmnet_meth$`signature plot`

### 5. Create a Data Frame with Genera and Coefficients. A data frame coda_balance_coeff_methis created to store genera names and their corresponding coefficients from the regression
coda_balance_coeff_meth <- data.frame(cbind(coda_glmnet_meth$`log-contrast coefficients`, coda_glmnet_meth$taxa.name))
View(coda_balance_coeff_meth)

### 6. Data Frame Manipulations
# Set row names of the data frame coda_balance_coeff_methto be the taxa names obtained from the regression results
rownames(coda_balance_coeff_meth) <- coda_balance_coeff_meth$taxa.name
# Set column names of the data frame to "Coefficient_full" and "Species" for clarity.
colnames(coda_balance_coeff_meth) <- c("Coefficient_full", "Species")
# Convert "Coefficient_full" column to numeric type, and create a new column "Coefficient" by rounding "Coefficient_full" to two decimal places
coda_balance_coeff_meth$Coefficient_full <- as.numeric(coda_balance_coeff_meth$Coefficient_full)
coda_balance_coeff_meth$Coefficient <-as.numeric(format(round(coda_balance_coeff_meth$Coefficient_full, 2), nsmall = 2))
# Determine Positive or Negative Coefficients. Create "PosNeg" column based on whether the coefficient is positive or negative; contains the labels "Positive" or "Negative" accordingly.
coda_balance_coeff_meth$PosNeg <- with(coda_balance_coeff_meth, ifelse(Coefficient > 0, "Positive", "Negative"))
# Assign Group Labels. The "Group" column is assigned labels "High" or "Low" based on the "PosNeg" column. If the coefficient is positive, it is labeled as "Low"; otherwise, it is labeled as "High".
coda_balance_coeff_meth$Group <- with(coda_balance_coeff_meth, ifelse(PosNeg == "Positive", "Low", "High"))
coda_balance_coeff_meth

### 7. Create Data Frame with Samples, Predictions, and Variables. A data frame coda_balances_meth_predictions is created to store balance predictions along with sex, group, and sample type information
coda_balances_meth_predictions <- data.frame(coda_glmnet_meth$predictions)
colnames(coda_balances_meth_predictions) <- c("Balance")
coda_balances_meth_predictions$CH4_level <- y_cu_vector_meth
# coda_balances_meth_predictions$Treatment <- factor(group[(group$Treatment=="Control"),]$Treatment)
coda_balances_meth_predictions$Treatment <- factor(group$Treatment)
# coda_balances_meth_predictions$Day <- factor(group[(group$Day=="Day 1" ),]$Day)
coda_balances_meth_predictions$Day <- factor(group$Day)
coda_balances_meth_predictions

### 8. Create Plots. The first one is a boxplot and scatter plot of microbial balances based on sex and group. The second one is a bar plot showing genera in the selected balance, filtered based on a coefficient threshold.
# Graph Sample Balances and Species Coefficients (coeff > 0.099)

# Species coefficient plot
#plot only the genera that are at least 10% of the contribution to the balance
coda_balance_coeff_meth_10p <- coda_balance_coeff_meth[(coda_balance_coeff_meth$Coefficient >= 0.099 | coda_balance_coeff_meth$Coefficient<= -0.099),]
coda_balance_coeff_meth_10p

####
#### simplify taxa names using simplify_taxa-names_count-tables.R
####

p_meth_sample_genera_meth<- ggplot(
  data = coda_balance_coeff_meth_10p, 
  aes(x = reorder(Species, Coefficient), y = Coefficient)) +
  geom_bar(stat="identity", width = 0.7, aes(color = Group, fill = Group), show.legend = FALSE) +
  scale_color_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  scale_fill_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  theme_classic() +
  labs(title = "Methane") +
  xlab(" ") + ylab("Log-contrast Coefficient") +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16, hjust = 0.5)) +
  coord_flip() 
p_meth_sample_genera_meth

# Balance prediction plot
p_meth_coda_sample_balances_meth<- ggplot(data = coda_balances_meth_predictions, aes(x = CH4_level, y = Balance, color = CH4_level)) +
  geom_line() +
#  geom_point() +
  geom_boxplot() +
  scale_color_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  scale_fill_manual(values = c("#F8766D","#00BFC4","#7CAE00","#C77CFF")) +
  theme_classic() +
  labs(title = "",
       color = "CH4 Level") +
  xlab(" ") + ylab("Balance Value") +
  #facet_grid(~Day) +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16)) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  #theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) +
  coord_flip() +
  theme(legend.position = "none")
p_meth_coda_sample_balances_meth


p_coda_glmnet_plots_meth <- egg::ggarrange(p_meth_sample_genera_meth, p_meth_coda_sample_balances_meth, labels = c("", ""), 
                                            ncol = 1, nrow = 2, heights = c(1,0.25))
as_ggplot(p_coda_glmnet_plots_meth)

```
Save plot
```{r}
ggsave(filename = "~/Documents/Programming/Microbiome/microbiome-ocp/data-analyses/results/coda4microbiome-pub/p_coda_glmnet_plots_meth.png", plot = p_coda_glmnet_plots_meth, width = 5, height = 4, dpi = 300)

```
### Run coda4microbiome algorithm
##### Microbial balances between Control and Users
```{r}
### Balances for Treatment Differences

### 1. Create a Transposed Data Frame. A transposed data frame x_species is created from the original data frame species.
x_species <- data.frame(t(species), check.names = T)
# subset asv/species/genus data to keep subgroup data
# x_species <- x_species[row.names(x_species) %in% group$SampleID, ]
x_species_control_oc <- x_species[row.names(x_species) %in% group_control_oc$SampleID, ]

### 2. Factorize the User/Control treatment groups. Create a factor variable cu_vector_control_oc by extracting treatment information for 'Control' & 'User' subjects from the y_control_oc(from group, from map) data.
y_vector_control_oc <- factor(group_control_oc$Treatment)

### 3. Apply regression. Regression is applied using the coda_glmnet function to identify important taxa in explaining CH4 level differences in 'Control' & 'User' subjects.
# only take taxa over 10% contribution to variance
coda_glmnet_control_oc <- coda_glmnet(x = x_species_control_oc, y = y_vector_control_oc, lambda = "lambda.min", nvar = NULL, alpha = 0.9, nfolds = 7, showPlots = TRUE, coef_threshold = 0)

### 4. Extract Accuracy Measure. The apparent Area Under the Curve (AUC) is extracted as an accuracy measure from the regression results.
coda_glmnet_control_oc$`apparent AUC`
coda_glmnet_control_oc$`predictions plot`
coda_glmnet_control_oc$`signature plot`

### 5. Create a Data Frame with Genera and Coefficients. A data frame coda_balance_coeff_control_ocis created to store genera names and their corresponding coefficients from the regression
coda_balance_coeff_control_oc <- data.frame(cbind(coda_glmnet_control_oc$`log-contrast coefficients`, coda_glmnet_control_oc$taxa.name))
# view(coda_balance_coeff_control_oc)

### 6. Data Frame Manipulations
# Set row names of the data frame coda_balance_coeff_control_octo be the taxa names obtained from the regression results
rownames(coda_balance_coeff_control_oc) <- coda_balance_coeff_control_oc$taxa.name
# Set column names of the data frame to "Coefficient_full" and "Species" for clarity.
colnames(coda_balance_coeff_control_oc) <- c("Coefficient_full", "Species")
# Convert "Coefficient_full" column to numeric type, and create a new column "Coefficient" by rounding "Coefficient_full" to two decimal places
coda_balance_coeff_control_oc$Coefficient_full <- as.numeric(coda_balance_coeff_control_oc$Coefficient_full)
coda_balance_coeff_control_oc$Coefficient <-as.numeric(format(round(coda_balance_coeff_control_oc$Coefficient_full, 2), nsmall = 2))
# Determine Positive or Negative Coefficients. Create "PosNeg" column based on whether the coefficient is positive or negative; contains the labels "Positive" or "Negative" accordingly.
coda_balance_coeff_control_oc$PosNeg <- with(coda_balance_coeff_control_oc, ifelse(Coefficient > 0, "Positive", "Negative"))
# Assign Group Labels. The "Group" column is assigned labels "High" or "Low" based on the "PosNeg" column. If the coefficient is positive, it is labeled as "Low"; otherwise, it is labeled as "High".
coda_balance_coeff_control_oc$Group <- with(coda_balance_coeff_control_oc, ifelse(PosNeg == "Positive", "Low", "High"))
coda_balance_coeff_control_oc

### 7. Create Data Frame with Samples, Predictions, and Variables. A data frame coda_balances_control_oc_predictions is created to store balance predictions along with sex, group, and sample type information
coda_balances_control_oc_predictions <- data.frame(coda_glmnet_control_oc$predictions)
colnames(coda_balances_control_oc_predictions) <- c("Balance")
coda_balances_control_oc_predictions$Treatment <- factor(group_control_oc$Treatment)
coda_balances_control_oc_predictions$Day <- factor(group_control_oc$Day)
coda_balances_control_oc_predictions

### 8. Create Plots. The first one is a boxplot and scatter plot of microbial balances based on sex and group. The second one is a bar plot showing genera in the selected balance, filtered based on a coefficient threshold.
# Graph Sample Balances and Species Coefficients (coeff > 0.099)

# Species coefficient plot
#plot only the genera that are at least 10% of the contribution to the balance
coda_balance_coeff_control_oc_10p <- coda_balance_coeff_control_oc[(coda_balance_coeff_control_oc$Coefficient >= 0.099 | coda_balance_coeff_control_oc$Coefficient<= -0.099),]
coda_balance_coeff_control_oc_10p

####
#### simplify taxa names using simplify_taxa-names_count-tables.R
####

coda_balance_coeff_control_oc_10p[7,2] <- "Eisenbergiella"

p_control_oc_sample_genera_control_oc<- ggplot(
  data = coda_balance_coeff_control_oc_10p, 
  aes(x = reorder(Species, Coefficient), y = Coefficient)) +
  geom_bar(stat="identity", width = 0.7, aes(color = Group, fill = Group), show.legend = FALSE) +
  scale_color_manual(values = c("#E6A316", "#E066A6","#7CAE00","#C77CFF")) +
  scale_fill_manual(values = c("#E6A316", "#E066A6","#7CAE00","#C77CFF")) +
  theme_classic() +
  labs(title = "Oral Contraceptives") +
  xlab(" ") + ylab("Log-contrast Coefficient") +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16, hjust = 0.5)) +
  coord_flip() 
p_control_oc_sample_genera_control_oc

# Balance prediction plot
p_control_oc_coda_sample_balances_control_oc<- ggplot(data = coda_balances_control_oc_predictions, aes(x = Treatment, y = Balance, color = Treatment)) +
  geom_line() +
#  geom_point() +
  geom_boxplot() +
  scale_color_manual(values = c("#E6A316", "#E066A6","#7CAE00","#C77CFF"),
                     labels = c("Control", "OC User", "", "")) +
  scale_fill_manual(values = c("#E6A316", "#E066A6","#7CAE00","#C77CFF")) +
  scale_x_discrete(labels=c("User" = "OC User")) +
  theme_classic() +
  labs(title = "",
       color = "Group") +  # Change legend title
  xlab(" ") + ylab("Balance Value") +
  #facet_grid(~Day) +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16)) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  #theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) +
  coord_flip() +
  theme(legend.position = "none")
p_control_oc_coda_sample_balances_control_oc

p_coda_glmnet_plots_control_oc<- egg::ggarrange(p_control_oc_sample_genera_control_oc, p_control_oc_coda_sample_balances_control_oc, labels = c("", ""), 
                                          ncol = 1, nrow = 2, heights = c(1,0.25))
as_ggplot(p_coda_glmnet_plots_control_oc)

```
Save plot
```{r}
ggsave(filename = "~/Documents/Programming/Microbiome/microbiome-ocp/data-analyses/results/coda4microbiome-pub/p_coda_glmnet_plots_control_oc.png", plot = p_coda_glmnet_plots_control_oc, width = 5, height = 4, dpi = 300)

```
### Run coda4microbiome algorithm
##### Microbial balances between *Outliers* and normal levels of H2 and CH4
```{r}
### Balances for Outliers vs. Normal Differences

### 1. Create a Transposed Data Frame. A transposed data frame x_species is created from the original data frame species.
x_species <- data.frame(t(species), check.names = T)
# subset asv/species/genus data to keep subgroup data
# x_species <- x_species[row.names(x_species) %in% group$SampleID, ]
x_species_out_norm <- x_species[row.names(x_species) %in% group_control_oc$SampleID, ]

### Create new data frame and add column 'Outliers' to categorize samples as Outlier or Normal (>print(outliers) for reference)
group_out_norm <- group_control_oc
group_out_norm$Outliers <- ifelse(group_out_norm$Subject %in% c(950, 962, 921, 735, 737), 'Outlier', 'Normal')
# check
group_out_norm

### 2. Factorize the Outlier/Normal levels of gas. Create a factor variable cu_vector_control_oc by extracting the methane information for 'Control' & 'User' subjects from the y_control_oc(from group, from map) data.
y_vector_out_norm <- factor(group_out_norm$Outliers)

### 3. Apply regression. Regression is applied using the coda_glmnet function to identify important taxa in explaining CH4 level differences in 'Control' & 'User' subjects.
# only take taxa over 10% contribution to variance
coda_glmnet_out_norm <- coda_glmnet(x = x_species_out_norm, y = y_vector_out_norm, lambda = "lambda.min", nvar = NULL, alpha = 0.9, nfolds = 6, showPlots = TRUE, coef_threshold = 0)

### 4. Extract Accuracy Measure. The apparent Area Under the Curve (AUC) is extracted as an accuracy measure from the regression results.
coda_glmnet_out_norm$`apparent AUC`
coda_glmnet_out_norm$`predictions plot`
coda_glmnet_out_norm$`signature plot`

### 5. Create a Data Frame with Genera and Coefficients. A data frame coda_balance_coeff_out_norm is created to store genera names and their corresponding coefficients from the regression
coda_balance_coeff_out_norm <- data.frame(cbind(coda_glmnet_out_norm$`log-contrast coefficients`, coda_glmnet_out_norm$taxa.name))
# View(coda_balance_coeff_out_norm)

### 6. Data Frame Manipulations
# Set row names of the data frame coda_balance_coeff_out_norm be the taxa names obtained from the regression results
rownames(coda_balance_coeff_out_norm) <- coda_balance_coeff_out_norm$taxa.name
# Set column names of the data frame to "Coefficient_full" and "Species" for clarity.
colnames(coda_balance_coeff_out_norm) <- c("Coefficient_full", "Species")
# Convert "Coefficient_full" column to numeric type, and create a new column "Coefficient" by rounding "Coefficient_full" to two decimal places
coda_balance_coeff_out_norm$Coefficient_full <- as.numeric(coda_balance_coeff_out_norm$Coefficient_full)
coda_balance_coeff_out_norm$Coefficient <-as.numeric(format(round(coda_balance_coeff_out_norm$Coefficient_full, 2), nsmall = 2))
# Determine Positive or Negative Coefficients. Create "PosNeg" column based on whether the coefficient is positive or negative; contains the labels "Positive" or "Negative" accordingly.
coda_balance_coeff_out_norm$PosNeg <- with(coda_balance_coeff_out_norm, ifelse(Coefficient > 0, "Positive", "Negative"))
# Assign Group Labels. The "Group" column is assigned labels "High" or "Low" based on the "PosNeg" column. If the coefficient is positive, it is labeled as "Low"; otherwise, it is labeled as "High".
coda_balance_coeff_out_norm$Group <- with(coda_balance_coeff_out_norm, ifelse(PosNeg == "Positive", "Low", "High"))
coda_balance_coeff_out_norm

### 7. Create Data Frame with Samples, Predictions, and Variables. A data frame coda_balances_control_oc_predictions is created to store balance predictions along with sex, group, and sample type information
coda_balances_out_norm_predictions <- data.frame(coda_glmnet_out_norm$predictions)
colnames(coda_balances_out_norm_predictions) <- c("Balance")
coda_balances_out_norm_predictions$Treatment <- factor(group_out_norm$Treatment)
coda_balances_out_norm_predictions$Day <- factor(group_out_norm$Day)
coda_balances_out_norm_predictions$Outliers <- factor(group_out_norm$Outliers)

### 8. Create Plots. The first one is a boxplot and scatter plot of microbial balances based on sex and group. The second one is a bar plot showing genera in the selected balance, filtered based on a coefficient threshold.
# Graph Sample Balances and Species Coefficients (coeff > 0.099)

# Species coefficient plot
#plot only the genera that are at least 10% of the contribution to the balance
coda_balance_coeff_out_norm_10p <- coda_balance_coeff_out_norm[(coda_balance_coeff_out_norm$Coefficient >= 0.099 | coda_balance_coeff_out_norm$Coefficient<= -0.099),]
coda_balance_coeff_out_norm_10p

####
#### simplify taxa names using simplify_taxa-names_count-tables.R
####

p_control_oc_sample_genera_out_norm <- ggplot(
  data = coda_balance_coeff_out_norm_10p, 
  aes(x = reorder(Species, Coefficient), y = Coefficient)) +
  geom_bar(stat="identity", width = 0.7, aes(color = Group, fill = Group), show.legend = FALSE) +
  scale_color_manual(values = c("#7CAE00","#C77CFF","#F8766D","#00BFC4")) +
  scale_fill_manual(values = c("#7CAE00","#C77CFF","#F8766D","#00BFC4")) +
  theme_classic() +
  labs(title = "Outliers") +
  xlab(" ") + ylab("Log-contrast Coefficient") +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16, hjust = 0.5)) +
  coord_flip() 
p_control_oc_sample_genera_out_norm

# Balance prediction plot
p_control_oc_coda_sample_balances_out_norm <- ggplot(data = coda_balances_out_norm_predictions, aes(x = Outliers, y = Balance, color = Outliers)) +
  geom_line() +
#  geom_point() +
  geom_boxplot() +
  scale_color_manual(values = c("#7CAE00","#C77CFF","#F8766D","#00BFC4"),
                     labels = c("Normal", "Outlier", "", "")) +
  scale_fill_manual(values = c("#7CAE00","#C77CFF","#F8766D","#00BFC4")) +
  # scale_x_discrete(labels=c("User" = "OC User")) +
  theme_classic() +
  labs(title = "",
       color = "Group") +  # Change legend title
  xlab(" ") + ylab("Balance Value") +
  #facet_grid(~Day) +
  theme(text = element_text(size = 15))  +
  theme(plot.title = element_text(size=16)) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  #theme(axis.text.y = element_text(angle = 90, hjust = 0.5)) +
  coord_flip() +
  theme(legend.position = "none")
p_control_oc_coda_sample_balances_out_norm

p_coda_glmnet_plots_out_norm <- egg::ggarrange(p_control_oc_sample_genera_out_norm, p_control_oc_coda_sample_balances_out_norm, labels = c("", ""), 
                                          ncol = 1, nrow = 2, heights = c(1,0.25))
as_ggplot(p_coda_glmnet_plots_out_norm)

```
Save plot
```{r}
ggsave(filename = "~/Documents/Programming/Microbiome/microbiome-ocp/data-analyses/results/coda4microbiome-pub/p_coda_glmnet_plots_out_norm.png", plot = p_coda_glmnet_plots_out_norm, width = 5, height = 4, dpi = 300)

```